### 2024-01-26 Paper/latex2/paper.tex
Add files via upload

%%
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%
%%
\documentclass[sigconf,screen,review,anonymous]{acmart}

\usepackage{multirow}
\usepackage{svg}
\usepackage{xcolor}

\newcommand{\todo}[1]{\textcolor{red}{\textbf{[TODO:#1]}}}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2024}
\acmYear{2024}
\acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[FSE '24]{The 32nd ACM Symposium on the Foundations of Software Engineering}{November 15--19, 2024}{Porto de Galinhas, Brazil}
%%
%%  Uncomment \acmBooktitle if the title of the proceedings is different
%%  from ``Proceedings of ...''!
%%
\acmBooktitle{Proceedings of the 32nd ACM Symposium on the Foundations of Software Engineering (FSE '24), November 15--19, 2024, Porto de Galinhas, Brazil}
\acmPrice{15.00}
\acmISBN{978-1-4503-XXXX-X/18/06}

\settopmatter{printacmref=false}
\renewcommand\footnotetextcopyrightpermission[1]{}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{RAX: A porting complexity evaluation tool for RISC-V}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Yuxuan Wang}
\email{wangyuxuan22@otcaix.iscas.ac.cn}
\affiliation{%
  \institution{Institute of Software, Chinese Academy of Sciences}
  \city{Haidian}
  \state{Beijing}
  \country{China}
}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Yuxuan et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  In recent years, the RISC-V’s open-source and reduced features have received active promotion of porting from the industry.Due to the lack of publicly available standards for assessing software portability, the software porting process for the RISC-V architecture has been predominantly reliant on expert experience, presenting obstacles. This paper introduces an automated tool, RAX, for assessing the porting complexity specifically for RISC-V by considering architecture-specific code and code Cyclomatic Complexity.RAX provides precise evaluation solutions and establishes a high-quality dataset for porting complexity assessment,enabling recommendations of suitable software difficulty levels for developers of varying skill levels.We continue to expand the dataset and conduct effective evaluations of the tool within the open-source community.We have gathered valuable feedback, including 47 positive responses, from 51 specific open-source projects.Interested users can explore the open-source implementation of RAX at (https://github.com/wangyuliu/R
  AX-2024), or watch the RAX video demo at (https://github.com/wan
  gyuliu/RAX-2024).
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
    <concept>
        <concept_id>10011007.10011006.10011072</concept_id>
        <concept_desc>Software and its engineering~Software libraries and repositories</concept_desc>
        <concept_significance>500</concept_significance>
        </concept>
    <concept>
        <concept_id>10011007.10011074.10011111</concept_id>
        <concept_desc>Software and its engineering~Software post-development issues</concept_desc>
        <concept_significance>300</concept_significance>
    </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Software libraries and repositories}
\ccsdesc[300]{Software and its engineering~Software post-development issues}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{ISA(Instruction set architecture), RISC-V, Porting Complexity Evaluation}

%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
%% \begin{teaserfigure}
%%  \includegraphics[width=\textwidth]{sampleteaser}
%%  \caption{Seattle Mariners at Spring Training, 2010.}
%%  \Description{Enjoying the baseball game from the third-base
%%  seats. Ichiro Suzuki preparing to bat.}
%%  \label{fig:teaser}
%% \end{teaserfigure}

%% \received{20 February 2007}
%% \received[revised]{12 March 2009}
%% \received[accepted]{5 June 2009}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}
\textcolor{orange}{In recent years, the RISC-V Instruction Set Architecture (ISA) has garnered significant attention from academia, particularly the industry, due to its open-source and reduced design.\citep{2014The}.OpenEuler has joined RISC-V International organization, and RISC-V has officially become openEuler's official support architecture.The mainline integration of openEuler RISC-V 23.09 was finished by the RISC-V SIG.The main repository of openEuler 23.03 has successfully ported 4,244 software packages to the RISCV64,while the openEuler 23.09 main repository has currently completed the porting work of 5,694 software packages\citep{osti_1560132}.OpenCloudOS Kernel Stream 2207.2 kernel version adds support for the RISC-V 64 architecture\citep{osti_1560133}.}

\textcolor{orange}{Despite the rapid development of the RISC-V ISA, which foundational software such as operating systems and compilers are relatively mature, there is still a significant demand for porting a large number of application software.\citep{2019Notary}.To adapt to the RISC-V,efficiently porting software package is greatly depending on domain expertise. During the initial planning phase of porting, the lack of standardized evaluation criteria for software portability can lead to misallocation of resources in the development process.Therefore, it is desirable to propose a method that automates the assessment of porting complexity,this will enable different development teams to access software packages that are tailored to their proficiency levels, thus assisting distribution vendors in the porting of RISC-V architecture.}

\textcolor{orange}{Du et al. have developed a basic statistical evaluation plan for CPU architecture code in project\citep{2023du},however,their work lacks a thorough validity analysis of factors,and fails to consider the complexity associated with the software operation and maintenance process.}

RAX can provide developers with accurate evaluation, recommend software with different levels of porting difficulty based on their skill levels. The contributions of this paper are as follows:
The contributions of this paper are as follows:
\begin{enumerate}
  \item In recent years, the RISC-V’s open-source and reduced features have received active promotion of porting from the industry.
  \item Optimization Assessment Tool.Tuning machine learning models to address ambiguous identification of samples of medium and hard complexity.
  \item We nearly doubled the dataset compared to other tool,and verified the effectiveness of the evaluation of 51 extension projects in real porting scenarios through their respective communities or forums.Our findings indicate that RAX outperforms other tool, delivering substantial improvements in terms of performance and accuracy.
\end{enumerate}

\section{RAX}
The following is an introduction to the process of RAX construction, as shown in Figure \ref{fig:figure1}.
During the analysis phase,architecture-related factors are obtained through the Commit tool for selection while investigating the impact of code complexity metrics on porting.
A scanning tool is built to get potential modification workload.
Finally,an appropriate machine learning model is selected as a classification.
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figure1.pdf}
  \caption{Flow chart for method to build RAX}
  \label{fig:figure1}
  \Description{Flow chart for method to build RAX}
\end{figure}
\subsection{Motivation}
To design RAX that assists developers with porting work,the following research questions are proposed:

\textbf{RQ 1:} How can we select architecture-relevant actors that is fine-grained to design an automated tool that assesses the porting difficulty?

\textbf{RQ 2:} How to select a classifier to obtain effective classification results for the tool?

To answer research question 1,this section provides a validity analysis of porting-related factors on RAX.
First,we try to use code complexity metrics to evaluate porting work.
The software metrics of code Cyclomatic Complexity(CC) can help developers make macro judgments on software complexity and maintenance difficulty, evaluate potential and hidden porting difficulties,which help the tool improve accuracy \cite{2005Exploring}.
At present, CC indicators are mostly used in project code quality assessment, defect inspection, code reconstruction, etc \cite{1991Cyclomatic}.
During the porting process, developers will target a large number of project source codes,build scripts,assembly codes,and a macro understanding of the project is required in the early stage of porting.
This part of the work has a lot of overlap with the scope of code complexity.Through our survey of community, we have discovered that even though architectural code can be separated quite clearly from common code, when faced with complex and large-scale projects, we still require numerous iterative tests to uncover unknown architectural code. Building upon this, we delve into the quantification of the significance of cyclomatic complexity in porting work.

\subsubsection{Factor 1: Cyclomatic Complex(CC)}
\label{sec:factor-1}
We uses the lizard CC detection tool and optimize its functions,and it will introduce it in Section \ref{sec:method}.
This study analyzes the relationship between CC and software porting complexity using the Spearman Rank correlation coefficient analysis method.
We invited five Master's students who are engaged in research on porting auxiliary tools to score the porting complexity(score) of 131 software adapted for RISC-V,collected from the OpenEule list\citep{stage2023} and GitHub.
According to the collection of Internet events, log analysis, source code dismantling, and similar project analogy \cite{liangguanyu2020}, we form a scoring interval in the range of $[0~5]$.
The Spearman coefficient of CC and score is 0.608, with a P-value of 1.39e-14, which is less than 0.01.
The results suggest that CC has a moderate positive correlation with porting difficulty.
This is because developers need a holistic understanding of the project during the porting process, and CC can represent the overall complexity of the project, potentially reflecting porting barriers, as shown in Figure \ref{fig:figure2}.

Second, we consider the effectiveness of the architecture binding code structure proposed by the other tool and conduct comparative experiments using more potential factors.
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figure2.pdf}
  \caption{The relationship between score and CC}
  \label{fig:figure2}
  \Description{The relationship between score and CC}
\end{figure}
\subsubsection{Factor2:Arch\_code}

Based on the analysis of previous evaluation tools, we selected factors with the greatest potential and that can be extracted through automated methods to define Arch\_code.
Arch\_code includes assembly code, conditional compilation and macro structure, Intrinsic functions, Builtin functions, architecture-related build scripts, and system calls.
We use an automated architecture binding code scanning tool to obtain statistics on the number of lines and frequencies of Arch\_code.

Continue to collect other factors that may affect the evaluation of porting complexity,and select the following parameters through the Commit analyzer:architecture-related code lines, architecture-related files, and architecture keywords.
The number of architecture-related code lines and files is located using the keywords of X86, ARM, MIPS, RISCV, SPARC and other architectures.
%Correlation comparison between Arch\_code and other factors acquired by Commit analyzer,as shown in Table \ref{tab:score}:

%\begin{table}
 % \caption{The Spearman correlation coefficient of score and architecture-related factors}
  %\label{tab:score}
  %\begin{tabular}{cc}
   % \toprule
    %Type & Score \\
    %\midrule
    %conditional compilation & 0.73 \\
    %Builtin & 0.61 \\
    %Intrinstic & 0.74 \\
    %syscall & 0.69 \\
    %asm & 0.66 \\
    %build\_scripts & 0.43 \\
    %time & -0.16 \\
    %authors & 0.43 \\
    %lines of arch & 0.16-0.40 \\
    %files of arch & 0.34-0.41 \\
    %keyword count & 0.17-0.43 \\
   % \bottomrule
  %\end{tabular}
%\end{table}
Correlation comparison between Arch\_code and other factors acquired by Commit analyzer,as shown in Figure \ref{fig:figure3}:
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figure3.pdf}
  \caption{The heat map of pearson correlation}
  \label{fig:figure3}
  \Description{The Correlation}
\end{figure}
The results indicate that Arch\_code has a higher correlation with score compared to other factors, and it also aligns with the significant results of the Peason correlation test.
The use of only architecture keywords for statistics yields moderate effectiveness in evaluating tool because keywords can only be detected from comments and macro sections of Commit, while critical porting code may lack these keywords,but the keyword count can reflect the amount of porting modification from the side.
As shown in the figure 2,the parameters of Arch\_code show a correlation of 0.45 to 0.75 with score, indicating a strong relationship.
However, using a single factor alone does not result in a significantly strong correlation, as the combined modifications of these factors make up the porting workload.

In the end,our tool selection was CC and the Arch\_code build scanning tool,as shown table 1.

% To design RAX that assists developers with porting work,the following research questions are proposed:

% \textbf{RQ 1:} How can we select architecture-relevant actors that is fine-grained to design an automated tool that assesses the porting difficulty?

% \textbf{RQ 2:} How to select a classifier to obtain effective classification results for the tool?

% \begin{table}
%   \caption{Tool factors comparison table}
%   \label{tab:comparisons}
%   \begin{tabular}{ccccc}
%     \toprule
%     Porting Complexity & \multicolumn{2}{c}{Du's Tool} & \multicolumn{2}{c}{RAX} \\
%      & TP & FN & TP & FN \\
%     \midrule
%     Low &32/37 & 5/37 & 35/37 & 2/37 \\
%     Medium & 5/8 & 3/8 & 6/8 & 2/8 \\
%     High & 2/6 & 4/6 & 6/6 & 0/6 \\
%     \midrule
%     Total & \textcolor{red}{39/51} & 12/51& \textcolor{red}{47/51} & 4/51 \\
%       \bottomrule
% \end{tabular}
% \end{table}

\begin{table}
  \caption{Tool factors comparison}
  \label{tab:Tool}
  \begin{tabular}{ccccccc}
    \toprule
    & Asm & Intrinsic & Builtin & Conditional compilation and macro & Buildscripts & CC \\
    \midrule
    Toolone & 1 & 1 & 1 & 1 & 1 & 0\\
    RAX & 1 & 1 & 1 & 1 & 1 & 1 \\
  \bottomrule
\end{tabular}
\end{table}

% \begin{table}
%   \caption{Tool factors comparison}
%   \label{tab:comparisont}
%   \begin{tabular}{ccccc}
%     \toprule
%      & Accuracy & Pmacro & Rmacro & F1macro \\
%     \midrule
%     Adaboost & 0.786 & 0.594 & 0.775 & 0.650 \\
%     Random Forest & \textcolor{red}{0.857} & 0.817 & 0.873 & 0.836 \\
%   \bottomrule
% \end{tabular}
% \end{table}

\subsection{Dataset}
We used the project upstream warehouse to build the data set,that projects mainly targeting C and C++ are from the OpenEuler\citep{stage2023} list and GitHub,and invited 5 postgraduate students who are engaged in the research of RISC-V porting auxiliary tools to participate in the experimental evaluation.Specific evaluation methods can be found in 2.1.1. The evaluation scores are divided into three categories: low, medium, and high\citep{githuburl}.This includes 139 projects as training, testing, and validation datasets, while more than 80 additional projects were used as a tool evaluation dataset. To gather insights from developers, the community, forums, and emails were utilized, resulting in over 51 meaningful feedback responses. In order to address the issue of data imbalance caused by collecting datasets solely from the openEuler mailing list, where more than 80\% of the projects were initially evaluated as having low porting complexity, a greater emphasis was placed on projects that had higher numbers of likes on GitHub and were deemed more important and active. Additionally, efforts were made to collect projects from the bottom such as operating systems, kernel-level projects,mathematical computation libraries,and machine learning libraries.

\subsection{Method}
\label{sec:method}
The RAX combines the CC calculation module and the Arch\_code location module to obtain the porting vector,and the complexity vector and score together form the data set.

\subsubsection{Lizard CC calculation scan function}
Lizard's default computing method is basically consistent with McCabe's theory \cite{1702388}.
However, Lizard tool only detects defects at the function level and uses the average function CC for file-level statistics, which is not suitable for current porting complexity detection requirements \cite{9402593}.
This paper proposes a new method for evaluating CC specifically for porting assessment work:
\begin{enumerate}
  \item First, calculate the function-level CC for all files in the project using McCabe's method.
  \item Following Liu's mention of System Cyclomatic Complexity (SCC) in software evolution assessment techniques based on code change detection \cite{liuhuihui00}, we define SCC as the total sum of function-level values in the system.
  The advantage is that it can avoid the use of overly complex methods, which will cause deviations to affect the classification results.
  Additionally,due to the system's complexity being calculated based on a single function,it eliminates the problem of decreased detection performance caused by the conversion of such a large code structure of all modules into into syntax abstract trees and control flow graphs.
\end{enumerate}

\subsubsection{Arch\_code location module}
we form the Arch\_code dictionary:
\begin{enumerate}
  \item Conditional compilation and architecture macros are used as preprocessor macros to enable projects to select execution on different hardware architectures, thus demonstrating the additional workload of architecture porting.
  Use conditional compilation statements, such as ``if defined'', ``ifdef'', in conjunction with all possible architectur macros,
  ``\_\_x86\_64\_\_''.
  We use a similar bracket matching algorithm to cover all the code within macros, avoiding the impact of nested layers.
  Additionally, we only collect the keyword dictionary for x86 architecture.
  This is because different architectures may reflect varying workloads and difficulties, while considering that multiple architectures blurs the classification performance in practical work.
  \item Assembly code Leverages the advantages of architecture-specific instructions to enhance performance.
  Considering that architecture changes inevitably require consideration of the assembly section, we use matching forms such as ``\_\_asm\_\_'' and ``\_\_volatile\_\_''.
  In future work,we will consider striving to achieve the currently unachievable architecture identification function of assembly code and identify assembly code in the form of instructions;In actual work, the assembly code of different architectures may represent different levels of difficulty.
  For example, the assembly code of ARM is relatively simple.
  At the same time,there are also situations where some assembly codes only need to serve a specific architecture and do not need to be fully modified when porting to RISC-V.
  This is why we cannot achieve Fine-grained evaluation,we simply count the number of lines.
  \item Intrinsic functions are interfaces provided by compilers that are closely related to the architecture.
  Calling Intrinsic functions can replace assembly code.
  We compile dictionaries of Intrinsic functions based on official documentation from x86.
  Builtin function also employ a similar approach;In the future,we may assign difficulty weights to them.
  Currently,we only perform frequency statistics on them.
  Additionally,we will continue to supplement the analysis of whether the functions already have corresponding implementation for RISC-V and whether they are truly needed in RISC-V.
  \item System calls are encapsulated by the operating system kernel to provide hardware calling.
  Considering glibc's support for system call encapsulation, reorganize the sys-call dictionary.
  The system calls that do not support RISC-V have been filtered out currently.
  \item In the porting process, adjustment of build scripts is necessary to ensure smooth compilation.
  We locate architecture keywords in the build scripts.
  In the future, architecture-related factors will consider even finer-grained classification.
  For instance,considering the impact of existing porting auxiliary tools, including intrinsic translation tools and assembly translation tools,on porting complexity assessment.
\end{enumerate}

\section{Result and analysis}

To answer research question 2, based on the work of Chen Xiang et al. \cite{5128907}, aiming at the status quo of unbalanced datasets and small sample datasets, we try to use ensemble learning methods \cite{6509481}, SMOTE \cite{4271036}, and random oversampling methods for data preprocessing.
Ultimately, we chose to train four models: Random Forest, Support Vector Machine, Adaboost, and Xgboost.
By comparing multiple indicators, we select the Random Forest machine learning as the classifier.

\begin{table}
  \caption{ Experimental results of models}
  \label{tab:evaluation}
  \begin{tabular}{ccccc}
    \toprule
     & Accuracy & Pmacro & Rmacro & F1macro \\
    \midrule
    Adaboost & 0.786 & 0.594 & 0.775 & 0.650 \\
    Random Forest & \textcolor{red}{0.857} & 0.817 & 0.873 & 0.836 \\
  \bottomrule
\end{tabular}
\end{table}

We continued to expand 100 projects as a supplementary data set, of which 87 were able to predict correctly.
The prediction results were determined through comments from the community and forums.
Appendix presents a comparative assessment of the porting complexity evaluation tool developed by Du Jiman et al. across various projects.Through real-world evaluations within the developer community, our tool has exhibited a significant improvement in accuracy, surpassing other comparable tools by a margin of 15.6\%. In the appendix, we provide a comprehensive compilation of the community's valuable feedback comments, accompanied by the relevant web links for further reference and validation.

\begin{table}
  \caption{Effectiveness comparison chart}
  \label{tab:effectiveness}
  \begin{tabular}{ccccc}
    \toprule
    Porting Complexity & \multicolumn{2}{c}{Du's Tool} & \multicolumn{2}{c}{RAX} \\
     & TP & FN & TP & FN \\
    \midrule
    Low &32/37 & 5/37 & 35/37 & 2/37 \\
    Medium & 5/8 & 3/8 & 6/8 & 2/8 \\
    High & 2/6 & 4/6 & 6/6 & 0/6 \\
    \midrule
    Total & \textcolor{red}{39/51} & 12/51& \textcolor{red}{47/51} & 4/51 \\
      \bottomrule
\end{tabular}
\end{table}

After adding the code complexity factor and modifying the original tool design model, RAX's result prediction accuracy is improved.
This is because code complexity can effectively distinguish projects with low porting complexity from projects with medium and high porting complexity.
Motivation Chapters present the cause analysis.
Adding factors also makes RAX more robust and more friendly to projects with unbalanced complexity vectors.
For example, kernel-level projects, boot layers, hardware drivers and other projects may involve a large amount of assembly code,corresponding to larger porting complexity,but other factors of the original tool design are at low values,which is not conducive to distinguishing complexity.
Chibios,u-boot are facing such a situation.
Within this assessment, there are instances of inaccurate predictions.

In-depth analysis of these variations reveals that certain projects exhibit inherent complexity, entailing a substantial presence of architecture-specific code defined by Arch\_code.
Nevertheless, the practical porting process benefits from an array of auxiliary tools designed to support developers. In response to valuable suggestions from developers within the OpenEuler community, we conducted an evaluation of the porting difficulty for various versions of the Chromium.Initially, we were able to accurately assess the porting difficulty, yielding promising results. However, when it came to evaluating the work involved in iterative updates between versions, existing tools proved insufficient for precise estimation.Sometimes developers design their work has generic code for unknown architecture and CPU architecture parts are well separated from the common code,that just need only to adjust the macro to recognize RISC-V,like nuttx.
\section{Related Work}
The development of the RISC-V ecosystem is rapidly progressing.
In the realm of scientific computing, Tine Blaise et al. \cite{osti_1830102} accomplished the porting of the OpenCL framework to RISC-V through a compiler framework, thereby expanding RISC-V's accessibility to a wider range of scientific computing applications.
In the context of architectural migration of foundational software and libraries, Cheng Yuanhu et al. \cite{2017Slow} designed an embedded multi-instruction set processor,addressing the software ecosystem challenges present in the RISC-V embedded domain.
As a result, auxiliary efforts in porting are highlighted as valuable endeavors.
This tool approaches code complexity from the perspective of its evaluation role in software evolution and maintenance,considering its influence on porting complexity assessment \cite{1993Software}.
Alvares et al. \cite{7844689} proposed using the development team's ``tolerance'' for code complexity to infer team development capabilities and performance indicators for projects.
Vard et al. \cite{2017Evaluating} emphasized that the internal quality of software impacts developers' capabilities and maintenance durations.
They conducted a comprehensive study on various existing code complexity metrics' validity and suggested incorporating empirically observed code features as complexity triggers.
\section{Conclusion}
We propose RAX, with the aim of recommending projects of varying porting difficulties to developers. We incorporate the code complexity to reflect developers' actual workload. We verified the effectiveness of the evaluation of 51 extension projects in real porting scenarios through their respective communities or forums. Observed significant performance improvements compared to previous tools.Moreover, RAX provides developers with the functionality to locate target code segments.Our next steps involve attempting text classification methods to categorize the text results of the scanning tool, designing more granular statistical methods for architecture-related factors.
%%
%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acks}
  This work was supported by the Strategic Priority Research Program of the Chinese Academy of Sciences (Grant No. XDA0320102) and Youth Innovation Promotion Association, Chinese Academy of Sciences (Grant No. 2023118).
\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{paper}

\end{document}