### 2024-01-29 Paper/latex2/paper.tex
Add files via upload

%%
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%
%%
\documentclass[sigconf,screen,review,anonymous]{acmart}

\usepackage{multirow}
\usepackage{svg}
\usepackage{xcolor}
\usepackage{hyperref}

\newcommand{\todo}[1]{\textcolor{red}{\textbf{[TODO:#1]}}}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2024}
\acmYear{2024}
\acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[FSE '24]{The 32nd ACM Symposium on the Foundations of Software Engineering}{November 15--19, 2024}{Porto de Galinhas, Brazil}
%%
%%  Uncomment \acmBooktitle if the title of the proceedings is different
%%  from ``Proceedings of ...''!
%%
\acmBooktitle{Proceedings of the 32nd ACM Symposium on the Foundations of Software Engineering (FSE '24), November 15--19, 2024, Porto de Galinhas, Brazil}
\acmPrice{15.00}
\acmISBN{978-1-4503-XXXX-X/18/06}

\settopmatter{printacmref=false}
\renewcommand\footnotetextcopyrightpermission[1]{}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{RAX: A porting complexity evaluation tool for RISC-V}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Yuxuan Wang}
\email{wangyuxuan22@otcaix.iscas.ac.cn}
\affiliation{%
  \institution{Institute of Software, Chinese Academy of Sciences}
  \city{Haidian}
  \state{Beijing}
  \country{China}
}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Yuxuan et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  In recent years, the open-source and reduced features of the RISC-V have led to active promotion of porting within the industry.Due to the lack of publicly available standards for assessing software portability, the software porting process for the RISC-V architecture has been predominantly reliant on expert experience, presenting obstacles. RAX is a automated tool which aids the C/C++ programmer in evaluating  complexity of projects during porting for RISC-V.RAX provides precise evaluation solutions and establishes a high-quality dataset for porting complexity assessment,enabling recommendations of suitable software with difficulty levels for developers of varying skill levels.We evaluate RAX on 72 popular and typical real-world C/C++ projects. The experimental result shows that 51 of them have been confirmed by developers within community.Interested users can explore the open-source implementation of RAX at \href{https://github.com/wangyuliu/RAX-2024}{https://github.com/wangyuliu/RAX-2024}, or watch the RAX video demo at \href{https://www.youtube.com/watch?v=SLphW3Ra4WM}{https://www.youtube.com/watch?v=SLphW3Ra4WM}.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
    <concept>
        <concept_id>10011007.10011006.10011072</concept_id>
        <concept_desc>Software and its engineering~Software libraries and repositories</concept_desc>
        <concept_significance>500</concept_significance>
        </concept>
    <concept>
        <concept_id>10011007.10011074.10011111</concept_id>
        <concept_desc>Software and its engineering~Software post-development issues</concept_desc>
        <concept_significance>300</concept_significance>
    </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Software libraries and repositories}
\ccsdesc[300]{Software and its engineering~Software post-development issues}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{ISA(Instruction set architecture), RISC-V, Porting Complexity Evaluation}

%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
%% \begin{teaserfigure}
%%  \includegraphics[width=\textwidth]{sampleteaser}
%%  \caption{Seattle Mariners at Spring Training, 2010.}
%%  \Description{Enjoying the baseball game from the third-base
%%  seats. Ichiro Suzuki preparing to bat.}
%%  \label{fig:teaser}
%% \end{teaserfigure}

%% \received{20 February 2007}
%% \received[revised]{12 March 2009}
%% \received[accepted]{5 June 2009}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}
In recent years, the RISC-V Instruction Set Architecture (ISA) has garnered significant attention from academia, and particularly the industry, due to its open-source and reduced design\citep{2014The}.OpenEuler has joined RISC-V International organization, and RISC-V has officially become openEuler's official support architecture.The mainline integration of openEuler RISC-V 23.09 was finished by the RISC-V SIG.The main repository of openEuler 23.03 has successfully ported 4,244 software packages to the RISCV64,while the openEuler 23.09 main repository has currently completed the porting work of 5,694 software packages\citep{osti_1560132}.OpenCloudOS Kernel Stream 2207.2 kernel version adds support for the RISC-V 64 architecture\citep{osti_1560133}.

Despite the rapid development of the RISC-V ISA, which foundational software such as operating systems and compilers are relatively mature, there is still a significant demand for porting a large number of application software\citep{2019Notary}.To adapt to the RISC-V,efficiently porting is greatly depending on domain expertise, and the lack of standardized evaluation criteria for software portability can lead to misallocation of resources.Therefore, it is desirable to propose a method that automates the assessment of porting complexity,this will enable different development teams to access software packages that are tailored to their proficiency levels, thus assisting distribution vendors in the porting of RISC-V architecture.

Many related works focus on the code complexity and portability evaluation\cite{2016helei},\cite{TAHIR2016101},\cite{Sholiq_2021}.Du et al. have developed a basic statistical evaluation plan for CPU architecture code\citep{2023du},however,their work lacks a thorough validity analysis of factors,and fails to consider the complexity factors associated with the software operation and maintenance process.
To overall these challenges, we propose RAX, a effective porting
complexity evaluation tool.The contributions of this paper are as follows:
\begin{enumerate}
  \item Identify one of the most important factors,Cyclomatic Complexity(CC), and investigate its significant value in porting scenarios.
  % \item We explore building high-quality datasets that can cover various porting scenarios to solve the problem of imbalanced datasets of existing tools.Tuning machine learning models to address ambiguous identification of samples of medium and hard complexity compared to other studies.
  \item We propose RAX, an automated tool to efficiently assess the complexity of software during porting to the RISC-V.
  % \item We nearly doubled the dataset compared to other tool,and verified the effectiveness of the evaluation of 51 extension projects in real porting scenarios through their respective communities or forums.Our findings indicate that RAX outperforms other tool, delivering substantial improvements in terms of accuracy.
  \item Verify the effectiveness of RAX in real scenarios concentrating C/C++ projects. Our findings indicate that RAX outperforms other tool, delivering substantial improvements in terms of accuracy.
\end{enumerate}

The remainder of this paper is organized as follows. Section2 analyze architecture-related factors briefly. Section 3 presents the implementation and evaluation of RAX. Section 4 and 5 presents related work and conclusions.

% \section{RAX}
% The following is an introduction to the process of RAX construction, as shown in Figure \ref{fig:figure1}.
% During the analysis phase,architecture-related factors are obtained through the Commit tool for selection while investigating the impact of code complexity metrics on porting.
% A scanning tool is built to get potential modification workload.
% Finally,an appropriate machine learning model is selected as a classification.
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figure1.pdf}
  \caption{Flow chart for method to build RAX}
  \label{fig:figure1}
  \Description{Flow chart for method to build RAX}
\end{figure}
\section{Motivation}
To design RAX that assists developers with porting work,the following research questions are proposed:

\textbf{RQ 1:} How can we select architecture-relevant actors that is fine-grained to design an automated assessment tool of porting complexity?

\textbf{RQ 2:} How to solve the problem of vague classification of projects with medium and high porting complexity in existing tools?
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figure2.pdf}
  \caption{The relationship between score and CC}
  \label{fig:figure2}
  \Description{The relationship between score and CC}
\end{figure}

To answer RQ 1,this section provides a validity analysis of porting-related factors on RAX.
First,we try to use code complexity metrics to evaluate porting work.
The software metrics of code Cyclomatic Complexity(CC) can help developers make macro judgments on software complexity and maintenance difficulty, evaluate potential and hidden porting difficulties,which help the tool improve accuracy \cite{2005Exploring}.
At present, CC indicators are mostly used in project code quality assessment, defect inspection, code reconstruction, etc \cite{1991Cyclomatic}.
During the porting process, developers will target a large number of project source codes,build scripts,assembly codes,and a macro understanding of the project is required in the early stage of porting.
This part of the work has a lot of overlap with the scope of code complexity.Through our survey of community, we discovered that even though architectural code can be separated quite clearly from common code, when faced with large-scale projects, we still require numerous iterative tests to uncover unknown architectural code. Building upon this, we delve into the quantification of the significance of cyclomatic complexity in porting work.
\subsection{Factor 1: Cyclomatic Complex(CC)}
\label{sec:factor-1}
We uses the lizard CC detection tool and optimize its functions,and it will introduce it in Section \ref{sec:Implementation}.
This study analyzes the relationship between CC and software porting complexity using the Spearman Rank correlation coefficient analysis method.
We invited five Master's students who are engaged in research on porting auxiliary tools to score the porting complexity(score) of 131 software adapted for RISC-V,collected from the OpenEuler list\citep{stage2023} and GitHub.
According to the collection of Internet events, log analysis, source code dismantling, and similar project analogy \cite{liangguanyu2020}, we form a scoring interval in the range of $[0,5]$.
The Spearman coefficient of CC and score is 0.608, with a P-value of 1.39e-14, which is less than 0.01.
The results suggest that CC has a moderate positive correlation with porting difficulty.
This is because developers need a holistic understanding of the project during the porting process, and CC can represent the overall complexity of the project, potentially reflecting porting barriers, the relationship of factors shown in Figure \ref{fig:figure2}.

Second, we consider the effectiveness of the architecture binding code structure proposed by the other tool and conduct comparative experiments using more potential factors.
\subsection{Factor2:Arch\_code}

We select factors with the greatest potential from existing tool and define Arch\_code.
Arch\_code includes assembly code, conditional compilation and macro structure(CCM), Intrinsic functions, Builtin functions, architecture-related build scripts, and system calls.
We use an automated tool to obtain statistics of lines and frequencies of Arch\_code.

Continue to collect other factors that may reflect the porting complexity,and select the following parameters through the Commit analyzer:architecture-related code lines, architecture-related files, and architecture keywords.architecture-related code lines and files is located using the keywords of the RISC-V architecture, such as riscv, rv32, rv64, RISC-V, RISCV and so on.
%Correlation comparison between Arch\_code and other factors acquired by Commit analyzer,as shown in Table \ref{tab:Tool}:

%\begin{table}
 % \caption{The Spearman correlation coefficient of score and architecture-related factors}
  %\label{tab:score}
  %\begin{tabular}{cc}
   % \toprule
    %Type & Score \\
    %\midrule
    %conditional compilation & 0.73 \\
    %Builtin & 0.61 \\
    %Intrinstic & 0.74 \\
    %syscall & 0.69 \\
    %asm & 0.66 \\
    %build\_scripts & 0.43 \\
    %time & -0.16 \\
    %authors & 0.43 \\
    %lines of arch & 0.16-0.40 \\
    %files of arch & 0.34-0.41 \\
    %keyword count & 0.17-0.43 \\
   % \bottomrule
  %\end{tabular}
%\end{table}

\begin{table*}
  \caption{Tool factors comparison}
  \label{tab:Tool}
  \begin{tabular}{ccccccccc}
    \toprule
    & Full Assembly & Inline Assembly & Intrinsic & Builtin & System Call & CCM & Buildscripts & CC \\
    \midrule
    Other tool & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \\
    RAX & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark \\
  \bottomrule
\end{tabular}
\end{table*}

Correlation comparison between Arch\_code,CC and other factors acquired by Commit analyzer,as shown in Figure \ref{fig:figure3}:
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figure3.pdf}
  \caption{The heat map of pearson correlation}
  \label{fig:figure3}
  \Description{The Correlation}
\end{figure}
The results indicate that Arch\_code and CC has a higher correlation with score compared to other factors, and we use Peason correlation method to build heat map.
The use of only architecture keywords for statistics yields moderate effectiveness because keywords can only be detected from comments and macro sections of Commit, while critical porting code may lack these keywords,but the keyword count can reflect the amount of porting modification from the side.
Arch\_code show a correlation of 0.45 to 0.75 with score, indicating a strong relationship.
However, using a single factor alone does not result in a significantly strong correlation, as the combined modifications of these factors make up the porting workload.

In the end,our tool selection was CC and the Arch\_code build scanning tool,as shown in Table \ref{tab:Tool}.

% To design RAX that assists developers with porting work,the following research questions are proposed:

% \textbf{RQ 1:} How can we select architecture-relevant actors that is fine-grained to design an automated tool that assesses the porting difficulty?

% \textbf{RQ 2:} How to select a classifier to obtain effective classification results for the tool?

% \begin{table}
%   \caption{Tool factors comparison table}
%   \label{tab:comparisons}
%   \begin{tabular}{ccccc}
%     \toprule
%     Porting Complexity & \multicolumn{2}{c}{Du's Tool} & \multicolumn{2}{c}{RAX} \\
%      & TP & FN & TP & FN \\
%     \midrule
%     Low &32/37 & 5/37 & 35/37 & 2/37 \\
%     Medium & 5/8 & 3/8 & 6/8 & 2/8 \\
%     High & 2/6 & 4/6 & 6/6 & 0/6 \\
%     \midrule
%     Total & \textcolor{red}{39/51} & 12/51& \textcolor{red}{47/51} & 4/51 \\
%       \bottomrule
% \end{tabular}
% \end{table}
% \documentclass{article}
% \usepackage{pifont} % 加载pifont宏包

% \begin{table*}
%   \caption{Tool factors comparison}
%   \label{tab:Tool}
%   \begin{tabular}{ccccccccc}
%     \toprule
%     & Full Assembly & Inline Assembly & Intrinsic & Builtin & System Call & Conditional compilation and macro & Buildscripts & CC \\
%     \midrule
%     Toolone & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \\
%     RAX & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark \\
%   \bottomrule
% \end{tabular}
% \end{table*}

% \begin{table*}
%   \caption{The Spearman correlation coefficient of score and architecture-related factors}
%   \label{tab:score}
%   \begin{tabular}{cccccc}
%     \toprule
%     conditional compilation & Builtin & Intrinstic & syscall & asm & build\_scripts \\
%     \midrule
%     0.73 & 0.61 & 0.74 & 0.69 & 0.66 & 0.43 \\
%   \bottomrule
% \end{tabular}
% \begin{tabular}{ccccc}
%   \toprule
%   time & authors & lines of arch & files of arch & keyword count \\
%   \midrule
%   -0.16 & 0.43 & 0.16-0.40 & 0.34-0.41 & 0.17-0.43 \\
% \bottomrule
% \end{tabular}
% \end{table*}

% \begin{table}
%   \caption{Tool factors comparison}
%   \label{tab:comparisont}
%   \begin{tabular}{ccccc}
%     \toprule
%      & Accuracy & Pmacro & Rmacro & F1macro \\
%     \midrule
%     Adaboost & 0.786 & 0.594 & 0.775 & 0.650 \\
%     Random Forest & \textcolor{red}{0.857} & 0.817 & 0.873 & 0.836 \\
%   \bottomrule
% \end{tabular}
% \end{table}

\section{RAX}
The following is an introduction to the process of RAX construction, as shown in Figure \ref{fig:figure1}.
During the analysis phase,architecture-related factors are obtained through the Commit tool for selection while investigating the impact of code complexity metrics on porting.
A scanning tool is built to get potential modification workload.
Finally,an appropriate machine learning model is selected as a classification.

% \subsection{Dataset}
% We used the project upstream warehouse to build the data set,that projects mainly targeting C and C++ are from the OpenEuler list and GitHub,and invited 5 postgraduate students who are engaged in the research of RISC-V porting auxiliary tools to participate in the experimental evaluation.Specific evaluation methods can be found in \ref{sec:factor-1}. The evaluation scores are divided into three categories: low, medium, and high\citep{githuburl}.This includes 139 projects as training, testing, and validation datasets, while more than 80 additional projects were used as a tool effectiveness evaluation dataset. To gather insights from developers, the community, forums, and emails were utilized, resulting in over 51 meaningful feedback responses. In order to address the issue of data imbalance caused by collecting datasets solely from the openEuler mailing list, where more than 80\% of the projects were initially evaluated as having low porting complexity, a greater emphasis was placed on projects that had higher numbers of likes on GitHub and were deemed more important and active. Additionally, efforts were made to collect projects from the system bottom such as operating systems, kernel-level projects,mathematical computation libraries,and machine learning libraries to address RQ 2.

\subsection{Implementation}
\label{sec:Implementation}
RAX obtains porting complexity vector through the CC calculation method and Arch\_code statistical value.

\subsubsection{Lizard CC calculation scan function}
Lizard's default computing method is basically consistent with McCabe's theory \cite{1702388}.
However, Lizard tool only defects at the function level and uses the average function CC for file-level statistics, which is not suitable for current porting complexity detection requirements \cite{9402593}.
This paper proposes a new method:

First, calculate the function-level CC for all files in the project using McCabe's method.Following Liu's mention of System Cyclomatic Complexity (SCC) in software evolution assessment techniques based on code change detection \cite{liuhuihui00}, we define SCC as the total sum of function-level values in the system.
  The advantage is that it can avoid the use of overly complex methods, which will cause deviations to affect the classification results.
  Additionally,due to the system's complexity being calculated based on a single function,it eliminates the problem of decreased detection performance caused by the conversion of such a large code structure of all modules into into syntax abstract trees and control flow graphs.
  Our CC calculation approach is expressed by the following equation,where lowercase cc represents the cyclomatic complexity at the function or file level.The uppercase CC represents the overall cyclomatic complexity of the software package,serving as a crucial factor in our tool.Here, |E| denotes the number of edges in the function control flow graph,and |N| represents the number of nodes.The function control flow graph is constructed based on branching structures such as if, while, and switch statements.

  \begin{equation}
    \mathrm{cc}(\mathrm{function}) = \lvert E \rvert - \lvert N \rvert + 2
    % \sum_{i=1}^{n} x_i
    \end{equation}
    \begin{equation}
      \mathrm{cc}(\mathrm{file}) =  \sum_{i=1}^{n} \mathrm{cc}(\mathrm{function})
      \end{equation}
      \begin{equation}
        \mathrm{CC} =  \sum_{i=1}^{k} \mathrm{cc}(\mathrm{file})
        \end{equation}

\subsubsection{Arch\_code scan function}
Other tools only consider x86 architecture to locate architecture-related code for statistical analysis. We propose a new idea that explore whether multiple architectures can be selected or if other architectures can be individually targeted. Based on this, we conducted comparative experiments using various architectures, including X86, ARM, MIPS, RISCV, SPARC, and PowerPC. We performed scanning and counting based on the keyword and function dictionaries specific to each architecture, which were derived from common patterns summarized from official websites.
By analyzing the correlation coefficient (Pearson's) between the keyword coverage line statistics of the six architectures and the score representing the porting complexity, we found that the correlation ranged from 0.44 to 0.58, indicating an average correlation across all architectures. Additionally, the correlation between the total values for all architectures and score was only 0.49.
Based on these results, analyze that our tool try to effectively assess the porting difficulty between two architectures. It is not as simple as adding up the modification amounts for different architectures, as it may only reflect an increase in code volume. Moreover, due to the large differences in row\_based statistical results caused by the differences in code structure reflected by each architecture, this approach that easily considers all architectures simultaneously without considering the code structure and architectural weight lead to ambiguous outcomes.Therefore, based on these experiments, we still select x86 to build Arch\_code scan function, given its mainstream position.

we form the Arch\_code dictionary:
\textbf{Conditional compilation and architecture macros} Use conditional compilation statements, such as ``if defined'', ``ifdef'', in conjunction with all possible architecture macros,
  ``\_\_x86\_64\_\_''.
  We use an algorithm similar to bracket matching to cover all the code within macros, avoiding the impact of nested layers.
  \textbf{Assembly} We use matching forms such as ``\_\_asm\_\_'' and ``\_\_volatile\_\_''.
  In future work,we will consider striving to achieve the currently unachievable architecture identification function of assembly code and identify assembly code in the form of instructions;
  \textbf{Intrinsic and Builtin} 
  We compile dictionaries of Intrinsic functions based on official documentation from x86\cite{x86intrin}.
  Builtin function also employ a similar approach;In the future,we may assign difficulty weights to them.
  \textbf{System calls} Considering glibc's support for system call encapsulation, reorganize the sys-call dictionary.
  The system calls that do not support RISC-V have been filtered and collected currently.
  \textbf{Buildscripts} We locate architecture keywords in the build scripts.

\subsection{Evaluation}
\subsubsection{Dataset}
We used the project upstream warehouse to build the data set,that projects mainly targeting C and C++ are from the OpenEuler list and GitHub,and invited 5 postgraduate students who are engaged in the research of RISC-V porting auxiliary tools to participate in the experimental evaluation.Specific evaluation methods can be found in \ref{sec:factor-1}. The evaluation scores are divided into three categories: low, medium, and high\citep{githuburl}.This includes 139 projects as training, testing, and validation datasets, while more than 80 additional projects were used as a tool effectiveness evaluation dataset. To gather insights from developers, the community, forums, and emails were utilized, resulting in over 51 meaningful feedback responses. In order to address the issue of data imbalance caused by collecting datasets solely from the openEuler mailing list, where more than 80\% of the projects were initially evaluated as having low porting complexity, a greater emphasis was placed on projects that had higher numbers of likes on GitHub and were deemed more important and active. Additionally, efforts were made to collect projects from the system bottom such as operating systems, kernel-level projects,mathematical computation libraries,and machine learning libraries to address RQ 2.
\subsubsection{Result and Analysis}
To answer RQ 2, based on the work of Chen Xiang et al. \cite{5128907}, aiming at the status quo of unbalanced datasets and small sample datasets, we try to use ensemble learning methods \cite{6509481}, SMOTE \cite{4271036}, and random oversampling methods for data preprocessing.
Ultimately, we chose to train four models: Random Forest, Support Vector Machine, Adaboost, and Xgboost.
By comparing multiple indicators, we select the Random Forest machine learning as the classifier, as shown in Table \ref{tab:evaluation}.

\begin{table}
  \caption{Experimental results of models}
  \label{tab:evaluation}
  \begin{tabular}{ccccc}
    \toprule
     & Accuracy & Pmacro & Rmacro & F1macro \\
    \midrule
    Adaboost & 0.786 & 0.594 & 0.775 & 0.650 \\
    Random Forest & \textcolor{red}{0.857} & 0.817 & 0.873 & 0.836 \\
  \bottomrule
\end{tabular}
\end{table}

We continued to expand 72 projects as a supplementary data set, and the prediction results were determined through comments from the community and forum.
Github presents comparison\cite{githubss} with the porting complexity evaluation tool developed by Du et al. across various projects, the comparison results are shown in Table \ref{tab:effectiveness}.Through real-world evaluations within the developer community, our tool has exhibited a significant improvement in accuracy, surpassing other comparable tools by a margin of 15.6\%. In Github, we provide a comprehensive compilation of the community's valuable feedback comments, accompanied by the relevant web links for further reference and validation.

\begin{table}
  \caption{Tool effectiveness comparison from community feedback}
  \label{tab:effectiveness}
  \begin{tabular}{ccccc}
    \toprule
    Porting Complexity & \multicolumn{2}{c}{Du's Tool} & \multicolumn{2}{c}{RAX} \\
     & TP & FN & TP & FN \\
    \midrule
    Low &32/37 & 5/37 & 35/37 & 2/37 \\
    Medium & 5/8 & 3/8 & 6/8 & 2/8 \\
    High & 2/6 & 4/6 & 6/6 & 0/6 \\
    \midrule
    Total & \textcolor{red}{39/51} & 12/51& \textcolor{red}{47/51} & 4/51 \\
      \bottomrule
\end{tabular}
\end{table}

After adding the code complexity factor and modifying the original model, RAX's result prediction accuracy is improved.
This is because code complexity can clearly distinguish three types of porting difficulty.
Motivation Chapters present the cause analysis.
Adding factors also makes RAX more robust and more friendly to projects with unbalanced complexity vectors.
For example, kernel-level projects, boot layers, hardware drivers and other projects may involve a large amount of assembly code,corresponding to larger porting complexity,but other factors of the original tool design are at low values,which is not conducive to distinguishing complexity.
Chibios,u-boot are facing such a situation. Another reason is that we have added more medium or difficult complex samples to the data set to increase robustness and improve the innovativeness of the tool.

Within assessment, there are instances of inaccurate predictions. Analysis of some variations reveals that certain projects exhibit inherent complexity, entailing a substantial presence of architecture-specific code defined by Arch\_code.
Nevertheless, the practical porting process benefits from an array of auxiliary tools designed to support developers. In response to valuable suggestions from developers within the OpenEuler community, we conducted an evaluation of the porting complexity for various versions of the Chromium.Initially, we were able to accurately assess the porting complexity, yielding promising results. However, when it came to evaluating the work involved in iterative updates between versions, existing tools proved insufficient for precise estimation.Sometimes developers design their work with generic code for unknown architecture and CPU architecture parts are well separated from the common code,that just need only to adjust the moderate amount of code to recognize RISC-V,like nuttx.
\section{Related Work}
The development of the RISC-V ecosystem is rapidly progressing.
In the realm of scientific computing, Tine Blaise et al. \cite{osti_1830102} accomplished the porting of the OpenCL framework to RISC-V through a compiler framework, thereby expanding RISC-V's accessibility to a wider range of scientific computing applications.
In the context of architectural porting of foundational software and libraries, Cheng Yuanhu et al. \cite{2017Slow} designed an embedded multi-instruction set processor,addressing the software ecosystem challenges present in the RISC-V embedded domain.
Therefore, auxiliary efforts in porting are highlighted as valuable endeavors.
Our work consider code complexity from the perspective of its evaluation role in software evolution and maintenance,considering its influence on porting complexity assessment \cite{1993Software}.
Alvares et al. \cite{7844689} proposed using the development team's ``tolerance'' for code complexity to infer team development capabilities and performance indicators for projects.
Vard et al. \cite{2017Evaluating} emphasized that the internal quality of software impacts developers' capabilities and maintenance duration.
They conducted a comprehensive study on various existing code complexity metrics' validity and suggested incorporating empirically observed code features as complexity triggers.
\section{Conclusion}
We propose RAX, with the aim of recommending projects of varying porting difficulties to developers. We incorporate the code complexity to reflect developers' actual workload. We verified the effectiveness of the evaluation of 51 extension projects in real porting scenarios through their respective communities or forums. Observed significant performance improvements compared to existing tools.Moreover, RAX provides developers with the functionality to locate target code segments.Our next steps involve attempting text classification methods to categorize the text results of the original tool, designing more granular statistical methods for architecture-related factors.
%%
%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acks}
  This work was supported by the Strategic Priority Research Program of the Chinese Academy of Sciences (Grant No. XDA0320102) and Youth Innovation Promotion Association, Chinese Academy of Sciences (Grant No. 2023118).
\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{paper}

\end{document}